### 2011-06-21 docs/introduction/introduction.tex
minor updates + add papers

MAESTRO describes the evolution of low Mach number flows that are in
hydrostatic equilibrium.  The idea for MAESTRO grew out of our success
in applying low Mach number combustion methods to small-scale
astrophysical flows.  Our original small-scale astrophysical
combustion algorithm is detailed in
\begin{itemize}
\item {\em Adaptive Low Mach Number Simulations of Nuclear Flames,}
J. B. Bell, M. S. Day, C. A. Rendleman, S. E. Woosley, \& M. Zingale
2004, JCP, 195, 2, 677 (henceforth BDRWZ)
\end{itemize}

@@

\noindent We have two papers that describe applications of the
method to Type Ia supernovae and X-ray bursts:
\begin{itemize}
\item {\em Multidimensional Modeling of Type I X-ray Bursts,}
  C. M. Malone, A. Nonaka, A. S. Almgren, J. B. Bell, \& M. Zingale 2011,
  ApJ, 728, 118 (henceforth ``the XRB paper'')
\item {\em The Convective Phase Preceding Type Ia Supernovae,}
  M. Zingale, A. S. Almgren, J. B. Bell, C. M. Malone, A. Nonaka, \& S. E. Woosley 2011,
  submitted to ApJ (henceforth paper V)
\end{itemize}

@@

In contrast, solving low Mach number systems (including the equations of
incompressible hydrodynamics) typically involves solving one or more
advection-like equations (representing, e.g.\ conservation of mass and
momentum) coupled with a divergence constraint on the velocity field.
For example, the equations of constant-density incompressible flow
are:
\begin{eqnarray}
U_t &=& -U \cdot \nabla U + \nabla p \label{eq:incompressible_u} \\
\nabla \cdot U &=& 0
\end{eqnarray}
Here, $U$ represents the velocity vector%
%
\footnote{Here we see an unfortunate conflict
of notation between the compressible hydro community and the
incompressible community.  In papers on compressible hydrodynamics,
$U$ will usually mean the vector of conserved quantities.  In 
incompressible / low speed papers, $U$ will mean the velocity vector.}
%
and $p$ is the dynamical pressure.  The time-evolution equation for
the velocity (eq.~[\ref{eq:incompressible_u}]) can be solved using
techniques similar to those developed for compressible hydrodynamics,
updating the old velocity, $U^n$, to the new time-level, $U^\star$.
Here the `$^\star$' indicates that the updated velocity does not, in
general, satisfy the divergence constraint.  A projection method will
take this updated velocity and force it to obey the constraint
equation.  The basic idea follows from the idea that any vector
field can be expressed as the sum of a divergence-free quantity and
the gradient of a scalar.  For the velocity, we can write:
\begin{equation}
U^\star = U^d + \nabla \phi \label{eq:decomposition}
\end{equation}
where $U^d$ is the divergence free portion of the velocity vector,
$U^\star$, and $\phi$ is a scalar.  Taking the divergence of
eq~(\ref{eq:decomposition}), we have
\begin{equation}
\nabla^2 \phi = \nabla \cdot U^\star
\end{equation}
(where we used $\nabla \cdot U^d = 0$).
With appropriate boundary conditions, this Poisson equation can be
solved for $\phi$, and the final, divergence free velocity can 
be computed as
\begin{equation}
U^{n+1} = U^\star - \nabla \phi
\end{equation}

>>>>

MAESTRO describes the evolution of low Mach number flows that are in
hydrostatic equilibrium.  The idea for MAESTRO grew out of our success
in applying low Mach number combustion methods developed for
terrestrial flames~\cite{DayBell00} to small-scale
astrophysical flows.  Our original small-scale astrophysical
combustion algorithm is detailed in
\begin{itemize}
\item {\em Adaptive Low Mach Number Simulations of Nuclear Flames,}
J. B. Bell, M. S. Day, C. A. Rendleman, S. E. Woosley, \& M. Zingale
2004, JCP, 195, 2, 677 (henceforth BDRWZ)
\end{itemize}

@@

\noindent We have several papers that describe applications of the
method to Type Ia supernovae and X-ray bursts:
\begin{itemize}
\item {\em Multidimensional Modeling of Type I X-ray Bursts,}
  C. M. Malone, A. Nonaka, A. S. Almgren, J. B. Bell, \& M. Zingale 2011,
  ApJ, 728, 118 (henceforth ``the XRB paper'')
\item {\em The Convective Phase Preceding Type Ia Supernovae,}
  M. Zingale, A. S. Almgren, J. B. Bell, C. M. Malone, A. Nonaka, \& S. E. Woosley 2011,
  submitted to ApJ (henceforth paper V)
\item {\em High-Resolution Simulations of Convection Preceding Ignition in
  Type Ia Supernovae Using Adaptive Mesh Refinement,}
  A. Nonaka, M. Zingale, A. J. Aspden, A. S. Almgren, J. B. Bell, \& S. E. Woosley
\end{itemize}

@@

In contrast, solving low Mach number systems (including the equations of
incompressible hydrodynamics) typically involves solving one or more
advection-like equations (representing, e.g.\ conservation of mass and
momentum) coupled with a divergence constraint on the velocity field.
For example, the equations of constant-density incompressible flow
are:
\begin{eqnarray}
U_t &=& -U \cdot \nabla U + \nabla p \label{eq:incompressible_u} \\
\nabla \cdot U &=& 0
\end{eqnarray}
Here, $U$ represents the velocity vector%
%
\footnote{Here we see an unfortunate conflict
of notation between the compressible hydro community and the
incompressible community.  In papers on compressible hydrodynamics,
$U$ will usually mean the vector of conserved quantities.  In 
incompressible / low speed papers, $U$ will mean the velocity vector.}
%
and $p$ is the dynamical pressure.  The time-evolution equation for
the velocity (Eq.~\ref{eq:incompressible_u}) can be solved using
techniques similar to those developed for compressible hydrodynamics,
updating the old velocity, $U^n$, to the new time-level, $U^\star$.
Here the `$^\star$' indicates that the updated velocity does not, in
general, satisfy the divergence constraint.  A projection method will
take this updated velocity and force it to obey the constraint
equation.  The basic idea follows from the fact that any vector
field can be expressed as the sum of a divergence-free quantity and
the gradient of a scalar.  For the velocity, we can write:
\begin{equation}
U^\star = U^d + \nabla \phi \label{eq:decomposition}
\end{equation}
where $U^d$ is the divergence free portion of the velocity vector,
$U^\star$, and $\phi$ is a scalar.  Taking the divergence of
Eq.~(\ref{eq:decomposition}), we have
\begin{equation}
\nabla^2 \phi = \nabla \cdot U^\star
\end{equation}
(where we used $\nabla \cdot U^d = 0$).
With appropriate boundary conditions, this Poisson equation can be
solved for $\phi$, and the final, divergence-free velocity can 
be computed as
\begin{equation}
U^{n+1} = U^\star - \nabla \phi
\end{equation}
